Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fate of human H3N2 influenza virus lineages

### 2018-01-15 paper/manuscript.tex
revisions to Abstract

\begin{abstract}
% no more than 250 words. (199 now)
The evolution of seasonal H3N2 influenza virus is mediated by the rapid accumulation of mutations. 
A better understanding of how mutations affect viral growth is critical for studying the pressures and constraints underlying the evolution of the influenza virus in nature. 
Here we experimentally measure the effect of every possible single amino-acid mutation to an H3 influenza hemagglutinin on viral growth in cell culture. 
We find the stalk domain to be fairly mutationally tolerant, and there is not a large disparity in tolerance between the head and stalk domains.
Furthermore, our measurements reveal mutations in successful seasonal influenza viral lineages to be significantly more preferred than mutations in lineages that have died out, and this was true at both epitope and non-epitope sites.
Epistasis in hemagglutinin is also seemingly driven by antigenic drift.
The preferences measured in a distantly related protein homolog do not reveal such differences in viral lineages.
A comparison of the preferences between the two hemagglutinin variants show substantial shifts in the effects of mutations.
Overall, our work highlights how experimental measurements of mutational effects can be leveraged to gain insight into the evolutionary fates of viral strains and potentially be used to predict viral evolution.
\end{abstract}

>>>>

\begin{abstract}
% no more than 250 words. (199 now)
Human influenza virus rapidly accumulates mutations in its major surface protein hemagglutinin (HA).
Forecasting the evolutionary success of influenza virus strains requires understanding how these mutations affect the protein's functionality and antigenicity.
Here we experimentally measure the effects on viral growth in cell culture of all single amino-acid mutations to the HA from a recent human H3N2 influenza virus strain. 
We show that these measurements are informative for understanding viral evolution in nature, since mutations in evolutionarily successful H3N2 viral lineages are measured to generally be more favorable than mutations in lineages that die out.
We also compare our measurements to prior data for an H1 HA, and find that the amino-acid preferences of many sites have substantially shifted between H1 and H3 HAs.
For instance, there is much less of a disparity in mutational tolerance between the head and stalk domains for the H3 HA.
Overall, our work suggests that experimental measurements of mutational effects might be leveraged to help forecast the evolutionary fates of viral lineages in nature.
\end{abstract}
